Testing Charts: viewer's perceptual accuracy in surveys

### 2023-07-20 SDSS-2023/testing-charts-accuracy.qmd
starting to restructure tasks 1-3

```{r data}
# other github repo. 
round1 <- readRDS(file.path(vs_repo_local, "round1.rds"))

@@

round2 <- readRDS(file.path(vs_repo_local, "round2.rds"))

@@

round3 <- readRDS(file.path(vs_repo_local, "round3.rds"))

# filter out all of the 'skipped on web':
round3_adj <- round3 %>% mutate(
  weight = WEIGHT_NOLA
) %>% filter(
  ab != "SKIPPED ON WEB",
  cd != "SKIPPED ON WEB"
)

round6 <- readRDS(file.path(vs_repo_local, "round6.rds"))

round6_adj <- round6 %>% mutate(
  weight = WEIGHT_NOLA
) %>% filter(RND_01 == 0) # same type of questions as round 7

round7 <- readRDS(file.path(vs_repo_local, "round7.rds"))

# filter out all of the 'skipped on web':
round7_adj <- round7 %>% mutate(
  weight = WEIGHT_NOLA
)
```

```{r data-rnd3}
round3 <- readRDS(file.path(vs_repo_local, "round3.rds"))

# filter out all of the 'skipped on web':
round3_adj <- round3 %>% mutate(
  weight = WEIGHT_NOLA
) %>% filter(
  ab != "SKIPPED ON WEB",
  cd != "SKIPPED ON WEB"
)
```

@@

We present viewers with structural variations on bar charts and pie charts and ask them to answer questions comparing the size of elements within those charts.
In this work, we present the series of tests we completed and the resulting findings.
The remainder of the paper is organized as follows: first, we describe the design of the visual stimulus used in our perception tests.
We then describe the population of study respondents and obtained survey sample.
Subsequently, we share our analyses of the resulting survey responses across each of our tests, including analyses on accuracy of responses and response behavior.
Finally, we discuss implications of this work and next steps.

@@

**How do we vary the task?** We ask participants to determine which of two just-noticeably different marked pieces is larger within the data visualization image, and we vary the structural design of that data visualization image.
We focus on three main sets of structural variation in the design.
First, we vary the alignment of the pieces in question.
Viewers are presented with two marked pieces in a chart that do not share a common baseline, then two pieces that do share a common baseline.
Second, we vary the orientation of the chart -- a vertically oriented stacked bar chart, a horizontally oriented version of the same chart, and then a wider version of the horizontally oriented chart.
\kr{We need to finalize what we end up showing in the results and then revisit this section...} Third, we consider a facetted bar chart and facetted pie chart.

@@

### Test 1: Alignment in stacked bar charts

@@

### Test 2: Orientation of stacked bar charts

In \autoref{fig-tasks-3} we see two structural variations on the stacked bar chart seen in Test 1: first, the Task 1 chart sized equally but rotated 90 degrees with bar lengths shown horizontally.
Second, the horizontal chart with a smaller aspect ratio so the bars appear thinner and longer.
Note that the wide version of the horizontal bars increases the pixel difference between the two marked pieces relative to the standard versions of the horizontal and vertical bars.

@@

### Test 3: Alignment and orientation in pie charts

@@

The data for this paper were collected in several rounds as part of the NORC Omnibus.

| Name    | Date  | # Participants | effective sample size | Sum of weights $\sum_i w_i$ |
|---------|-------|--------:|----------:|----------:|
| Round 1 | April 2022 |  `r nrow(round1_adj)`    |    `r round(neff(round1_adj$weight),1)`       | `r round(sum(round1_adj$weight),1)`       |
| Round 2 | May 2022 |  `r nrow(round2_adj)`   |    `r round(neff(round2_adj$weight),1)`       | `r round(sum(round2_adj$weight),1)`       |
| Round 3 | Jun 2022 |  `r nrow(round3_adj)`  |    `r round(neff(round3_adj$weight),1)`       | `r round(sum(round3_adj$weight),1)`       |
| Round 6 | Sep 2022 |  `r nrow(round6_adj)` [Split]  |    `r round(neff(round6_adj$weight),1)`       | `r round(sum(round6_adj$weight),1)`       |
| Round 7 | Oct 2022 |  `r nrow(round7_adj)`   |    `r round(neff(round7_adj$weight),1)`       | `r round(sum(round7_adj$weight),1)`       |
|         |       |         |           |           |

: Survey rounds: dates, number of participants (nominal sample size), effective sample size, and sum of weights. {#tbl-rounds}

@@

```{r combine_rounds_1_and_2}

lambda <- calculate_lambda(round1_adj$weight, round2_adj$weight)
round_12 <- round1_adj %>% 
            mutate(weight = lambda[1]*weight) %>% 
            select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD, weight) %>%
    rbind(round2_adj %>% 
            mutate(weight = (1-lambda[1])*weight) %>% 
            select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD, weight))
round_12_long <- round_12 %>% 
  pivot_longer(CorrectAB:CorrectCD, names_to = "Question", values_to = "Response")

>>>>

```{r data}
# other github repo. 
round1 <- readRDS(file.path(vs_repo_local, "round1.rds"))
round2 <- readRDS(file.path(vs_repo_local, "round2.rds"))

@@

@@

survey1 <- survey::svydesign(
  data = round1_adj, 
  ids = ~CaseId, weights = ~weight)

survey2 <- survey::svydesign(
  data = round2_adj,
  ids = ~CaseId, weights = ~weight)
```

@@

We present viewers with structural variations on bar charts and pie charts and ask them to answer questions comparing the size of elements within those charts.
In this work, we present the series of tests we completed and the resulting findings.
The remainder of the paper is organized as follows: first, we describe the design of the visual stimuli used in our perception tests.
We then describe the population of study respondents and obtained survey sample.
Subsequently, we share our analyses of the resulting survey responses across each of our tests, including analyses on accuracy of responses and response behavior.
Finally, we discuss implications of this work and next steps.

@@

**How do we vary the task?** We ask participants to determine which of two just-noticeably different marked pieces is larger within the data visualization image, and we vary the structural design of that data visualization image.
We focus on three main sets of structural variation in the design.
First, we vary the alignment of the pieces in question.
Viewers are presented with two marked pieces in a chart that do not share a common baseline, then two pieces that do share a common baseline.
Second, we vary the orientation of the chart -- we rotate the vertical stacked bar chart, and present a horizontally oriented version of the same chart, with identically sized marked pieces. Finally, we change the aspect ratio of the chart and present a wider version of the horizontally oriented chart which has longer marked pieces.

@@

### Task 1: Alignment in stacked bar charts

@@

### Task 2: Orientation of stacked bar charts

Story here: not a huge difference, but we mainly see a difference in how people react. horizontal/vertical/horizontal. Even if there's not hugely significant differences by design, we can see differences in behavior.

Combine Tasks 1 and 2 and just present overall, aligned/unaligned by the 3 different orientations? Doesn't matter whether we have a vertical or horizontal position,

We can reproduce
We can rank
We can rank better because we have a bigger sample size
WE can find changes in viewer behavior
Demographics matter to accuracy and choices

In \autoref{fig-tasks-3} we see two structural variations on the stacked bar chart seen in Task 1: first, the Task 1 chart sized equally but rotated 90 degrees with bar lengths shown horizontally.
Second, the horizontal chart with a smaller aspect ratio so the bars appear thinner and longer.
Note that the wide version of the horizontal bars increases the pixel difference between the two marked pieces relative to the standard versions of the horizontal and vertical bars.

@@

### Task 3: Alignment and orientation in pie charts

@@

@@

```{r lambdas}
n1 <- nrow(round1_adj)
n2 <- nrow(round2_adj)
neff1 <- svytotal(~one, survey1, deff=TRUE)
neff2 <- svytotal(~one, survey2, deff=TRUE)
#abcd1 <- svytotal(~CorrectABCD, survey1, deff=TRUE)
#abcd2 <- svytotal(~CorrectABCD, survey2, deff=TRUE)

d1 <- 1 + cv(neff1)^2
d2 <- 1 + cv(neff2)^2

lambda <- n1/(d1*denom) 
```

```{r combine_rounds_1_and_2}
round_12 <- round1_adj %>% 
            mutate(weight = lambda[1]*weight) %>% 
            select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD, weight) %>%
    rbind(round2_adj %>% 
            mutate(weight = (1-lambda[1])*weight) %>% 
            select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD, weight))
round_12_long <- round_12 %>% 
  pivot_longer(CorrectAB:CorrectCD, names_to = "Question", values_to = "Response")
